---
title: "README"
author: ""
date: ""
output: github_document
--- 

# Exponential Families

Zachary McCaw <br>
Updated: 23-01-07

```{r global_options, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

### Description

Collection of functions for working with exponential family distributions of the form:
$$
f(x) = \exp\left\{\sum_{k=0}^{K}\beta_{k}x^{k}\right\}
$$

Notes: 

* If $K$ is even, the support of $f(x)$ is assumed to be $(-\infty, \infty)$. If $K$ is odd, the support is assumed to be $(0, \infty)$. 

* The coefficient of the highest order term, i.e. $\beta_{K}$ is assumed to be negative. 

* The coefficient of the constant term, i.e. $\beta_{0}$ is the log of the normalizing constant.


## Installation

```{R, eval=FALSE}
devtools::install_github(repo = "zrmacc/ExpFam")
```


## Functions

### Normalizing constant

Given the coefficients of the non-constant terms $(\beta_{K}, \beta_{K-1}, \dots, \beta_{1})$, determine the log normalizing constant $\beta_{0}$. Note the coefficients of the non-constant terms are given in descending order. 

```{r}
library(ExpFam)
# Example of the standard normal distribution.
b_in <- c(-0.5, 0)
b_out <- NormalizingConstant(b_in)
print(b_out, digits = 3)
all.equal(b_out[3], -0.5 * log(2 * pi), tolerance = 0.0001)
```


### Density

Obtain the density function. Note that the normalizing constant is provided hereafter.

```{r}
# Example of the standard normal distribution.
b <- c(-0.5, 0, -0.5 * log(2 * pi))
dens <- GetDensity(b)
print(dens(0), digits = 3)
all.equal(dens(0), dnorm(0), tolerance = 0.0001)
```


### Cumulative distribution

Obtain the cumulative distribution function.

```{r}
# Example of the standard normal distribution.
b <- c(-0.5, 0, -0.5 * log(2 * pi))
cdf <- GetCDF(b)
print(cdf(1), digits = 3)
all.equal(cdf(1), pnorm(1), tolerance = 0.0001)
```


### Quantile

Evaluate the quantile. Note `GetQuantile` does not return a function but instead performs the quantile evaluation. 

```{r}
# Example of the standard normal distribution.
b <- c(-0.5, 0, -0.5 * log(2 * pi))
quants <- GetQuantile(b, probs = c(0.25, 0.50, 0.75))
print(quants, digits = 3)
all.equal(quants, qnorm(p = c(0.25, 0.50, 0.75)), tolerance = 0.0001)
```


### Sampling

Sample from an exponential family distribution. The utility of `rExpFam` is the ability to sample from distributions with K > 2, for which built-in functions (e.g. `rnorm`) are not available. The sample is obtained by numerically inverting the CDF. `rExpFam` requires an approximate search interval. To determine this interval, it is helpful to plot the density. 

```{r}
b_in <- c(-1, 2.5, -2.1875, 0.78125)
b <- NormalizingConstant(b_in)
```

Density being sampled: 

```{r, cache=TRUE, fig.align='center'}
library(ggplot2)

# Plot density.
dens <- GetDensity(b)
x <- seq(from = -2, to = 3, length.out = 501)
df <- data.frame(x = x, y = dens(x))
q <- ggplot(data = df) + 
  theme_bw() + 
  geom_line(
    aes(x = x, y = y),
    linewidth = 1.2,
    color = "#0072B5FF"
  ) + 
  labs(
    x = "Value",
    y = "Density"
  ) 
show(q)
```

```{r}
x <- rExpFam(b = b, n = 10, xmin = -2, xmax = 3)
round(x, digits = 3)
```
